---
title: "Untitled"
format: html
editor: visual
---

```{r}
library(dplyr)
library(readxl) 
library(stringr)
library(rio)
library(here)
library(janitor)
library(tidyr)
```

# Reason data field fullname & shortname by document source

```{r}
#gasb
gasb_fullname <- import("data/HGarb_Updates_2022/Texas_ESRS_updated_2022.xlsx", sheet = "GASB 68") %>% colnames()

gasb_shortname <- import("data/HGarb_Updates_2022/Texas_ESRS_updated_2022.xlsx", sheet = "GASB 68", skip = 1) %>% clean_names() %>% 
  setnames(
    old = c("adec_amt", "adec_paid_amt", "adec_missed"),
    new = c(paste0(c("adec_amt", "adec_paid_amt", "adec_missed"), "_from_gasb"))) %>% 
  colnames() 

#valuation
valuation_fullname <- import("data/HGarb_Updates_2022/Texas_ESRS_updated_2022.xlsx", sheet = "Actuarial Valuation") %>% colnames()

valuation_shortname <- import("data/HGarb_Updates_2022/Texas_ESRS_updated_2022.xlsx", sheet = "Actuarial Valuation", skip = 1) %>% clean_names() %>% 
  # Differentiate some cols from sheet 2 & sheet 3 
      setnames(
        old =  c("actuarial_return", "market_return","ava" , "aal","ual", "funded_ratio_old", "payroll"), 
        new = c(paste0(c("actuarial_return", "market_return","ava" , "aal", "ual", "funded_ratio_old", "payroll"), "_from_valuation"))) %>% 
  colnames()
  
#acfr
acfr_fullname <- import("data/HGarb_Updates_2022/Texas_ESRS_updated_2022.xlsx", sheet = "CAFR") %>% colnames()

acfr_shortname <- import("data/HGarb_Updates_2022/Texas_ESRS_updated_2022.xlsx", sheet = "CAFR", skip = 1) %>% clean_names() %>%  
  # Differentiate some cols from sheet 2 & sheet 3 
   setnames(old = c("actuarial_return", "market_return", "ava", "aal", "ual", "funded_ratio_old",
"payroll"),
    new = c(paste0(c("actuarial_return", "market_return", "ava", "aal", "ual", "funded_ratio_old",
"payroll"), "_from_acfr"))) %>% 
  
  # Differentiate sheet 1 & sheet 3
  setnames(
    old = c("adec_amt", "adec_paid_amt", "adec_missed"),
    new = c(paste0(c("adec_amt", "adec_paid_amt", "adec_missed"), "_from_acfr"))) %>% 
  colnames()

data.frame(field_name_full = c(gasb_fullname, valuation_fullname, acfr_fullname),
           field_name_short = c(gasb_shortname, valuation_shortname, acfr_shortname)
  )

reason_field_name_full_short_source <- (
   # collect all names in gasb
          data.frame(field_name_full = gasb_fullname,
          field_name_short = gasb_shortname) %>% 
          mutate(source = "gasb")) %>% 
  
  # bind with valuation
          rbind(data.frame(field_name_full = valuation_fullname, 
          field_name_short = valuation_shortname) %>% 
          mutate(source = "valuation")) %>% 
          slice(-c(1:2)) %>% #remove full_name and fye
  
  # bind with acfr
         rbind(data.frame(field_name_full = acfr_fullname, 
         field_name_short = acfr_shortname) %>% 
         mutate(source = "acfr")) %>% 
         slice(-c(1:2))  #remove full_name and fye

write.csv(reason_field_name_full_short_source, "output/reason_full_short_fieldname_by_document.csv")
```

# PPD code book & field names

\`\`\`{r} \# downloaded from ppd site ppd_colnames \<- import(here("data/ppd-data-latest.csv")) %\>% colnames() %\>% tolower()%\>% sort()

#https://publicplansdata.org/public-plans-database/download-full-data-set/ ppd_colnames_from_codebook \<- import(here("data/ppd-codebook.csv")) %\>% clean_names() %\>% \# change rename(ppd_fullname = variable_label) %\>% mutate(ppd_shortname = str_to_lower(variable_name)) %\>% select(ppd_fullname, ppd_shortname)

# it seems like Reason has slightly different ppd colnames:

ppd_colnames_created_by_reason \<-import(here::here("data/marc_swaroop_handover_files/ppd-out.xlsx"))%\>% colnames() %\>% sort()

write.csv(ppd_colnames_from_codebook, "output/ppd_colnames_from_codebook.csv")
